Ionization of biological molecules by multicharged ions using the stoichiometric model

### 2019-09-28 arxiv/stoi_ionization.tex
figs and text modified with 27/09 corrections

%\documentclass[pra,showpacs,twocolumn]{revtex4}
\documentclass[preprint,showpacs]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\title{Ionization of biological molecules by multicharged ions by using 
the stoichiometric model}

\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
%\author{author}
%\author{author}
\address{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET--UBA), \\ Buenos Aires, Argentina.}

\date{\today}% It is always \today, today,
\pacs{34.50Gb, 34.80Gs, 34.80Dp}

\begin{abstract}
In the present work, we investigate the ionization of molecules of 
biological interest by the impact of multicharged ions in the 
intermediate to high energy range. We performed full non--perturbative 
distorted--wave calculations (CDW) for thirty--six collisional systems. 
Six atomic targets compose these systems: H, C, N, O, F and S --which 
are the constituents of most of the DNA and biological molecules-- and 
six charged projectiles (antiprotons, H, He, B, C, and O). On account 
of the radiation damage caused by secondary electrons, we also inspect 
the energy and angular distributions of the emitted electrons from the 
atomic targets. We examine seventeen molecules: DNA and RNA bases, 
DNA backbone, pyrimidines, tetrahydrofuran (THF), and several 
C$_n$H$_n$ compounds. We show that the simple stoichiometric model 
(SSM), which approximates the molecular ionization cross sections as a 
linear combination of the atomic ones, gives reasonably good results 
for complex molecules. We also inspect the extensively used Toburen 
scaling of the total ionization cross sections of molecules with the 
number of weakly bound electrons. Based on the present atomic CDW 
results, we propose new active electron numbers. The CDW--based scaling 
improves significantly the single curve behavior for all the targets 
and ions studied here, from the intermediate to the high energy region. 
Furthermore, the new scaling describes well the available experimental 
data for proton impact, including small molecules such as H$_2$, water, 
methane, and ammonia. For completeness, we perform full molecular 
calculations for five nucleobases and test a modified stoichiometric 
formula based on the Mulliken charge of the composite atoms. The 
difference introduced by the new stoichiometric formula is less than 
$3\%$, which indicates the reliability and robustness of the SSM to 
deal with this type of molecules. The results of the extensive 
ion--target examination included in the present study allow us to 
assert that the SSM and the CDW--based scaling will be useful tools in 
this area.
\end{abstract}

%\keywords{Suggested keywords}

\maketitle

%\tableofcontents
%\pacs{79.20.Rf, 68.49.Bc, 34.20.Cf,34.50.-s,34.35.+a}

@@

In Section~\ref{subsec:stoichiometric}, we deal with the molecular 
structure complexity of the targets by 
implementing the simplest stoichiometric model (SSM): the molecules are 
assumed to be composed of isolated independent atoms, and the total 
cross section by a linear combination of stoichiometric weighted atomic 
calculations.
By implementing the CDW and the SSM, we calculate ionization 
cross section of several molecules of biological interest, including 
DNA and RNA molecules, such as adenine, cytosine, guanine, thymine, uracil, 
tetrahydrofuran~(THF), pyrimidine, and DNA backbone, by the impact 
of antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
In Section~\ref{subsec:scaling}, we
test the Toburen scaling rule~\cite{toburen1975,toburen1976}, which 
states that the ratio between the ionization cross section and the 
number of weakly bound electrons can be arranged in a narrow universal 
band in terms of the projectile velocity. We applied this rule to 
several hydrocarbons and nucleobases and noted that the width of the 
resulting universal band could be significantly reduced if we consider 
the number of active electrons in the collision based on the 
CDW results for the different atoms. The new scaling was then tested 
theoretically and by comparison with experimental data available.

@@

In the present study, we consider six atoms, $\alpha=$ H, C, N, O, P, 
and S, and six projectiles, antiprotons $\bar{p}$, H$^{+}$, He$^{+2}$, 
Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
Most of the organic molecules are composed of these atoms. Some 
particular molecules also include halogen atoms such as fluor and 
bromine; ionization cross sections of these elements have been 
previously published~\cite{miraglia2008}.

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig1.eps}
\includegraphics[width=0.8\textwidth]{figuras/atomicscaling.eps}
\caption{Reduced CDW total ionization cross section of six atomic 
targets. The curves are labeled with the charge state corresponding to 
the six multicharged projectiles.}
\label{fig:atomscaling}
\end{figure*}

We display our total CDW ionization cross sections for the six essential 
elements by the impact of the six projectiles in Fig.~\ref{fig:atomscaling}.
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross section scales with the square of the projectile 
charge, $Z^{2}$. The impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 keV. We also performed similar calculations with the 
first Born approximation, and we corroborated that it provides quite 
reliable results only for energies higher than a couple of MeV/amu. 
We use the same line color to indicate the projectile charge throughout 
all the figures of this work: dashed--red, solid--red, blue, magenta, 
olive and orange for antiprotons, H$^{+}$, He$^{+2}$, Be$^{+4}$, 
C$^{+6}$, and O$^{+8}$, respectively. Notably, there is no complete 
tabulation of ionization of atoms by the impact of multicharged ions. 
We hope that the ones presented in this article will be of help for 
future works.

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig5.eps}
\includegraphics[width=0.8\textwidth]{figuras/ener_mean.eps}
\caption{Mean emitted energy distribution for ionization by the impact 
of multicharged ions, given by Eq.~(\ref{40}). Dashed lines for $\bar{p}$,
solid lines for ion charges $+1$, $+2$, $+4$, $+6$ and $+8$,
as indicated.}
\label{fig:emittedener}
\end{figure*}

@@

In Fig.~\ref{fig:emittedener}, we estimate $\overline{E}_{\alpha}$ of
the emitted electron in the 0.5--2.7 a.u. energy range, or equivalently 
from 15 to 70 eV, for all the targets. Our results agree with the 
experimental findings~\cite{surdutovic2018}. As can be noted in the 
figure, the mean energy value is surprisingly sensitive to the 
projectile charge $Z$, which can duplicate the proton results in the 
intermediate region, i.e., 100--400 keV/amu. The effect observed can be 
attributed to the depletion caused by the multicharged ions to the 
yields of low energy electrons. This behavior cannot be found in the 
first Born approximation, where the $Z^2$ law cancels the $Z$ dependence
in Eq.~(\ref{40}). At high energies, $\overline{E}_{\alpha}$ tends to a 
universal value for all ions, as can be seen in Fig.~\ref{fig:emittedener}.

@@

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig6.eps}
\includegraphics[width=0.8\textwidth]{figuras/ang_mean.eps}
\caption{Mean emitted angle distribution for ionization by impact of
multicharged ions.}
\label{fig:emittedang}
\end{figure*}

@@

It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons is nearly isotropic. This 
behavior is caused by the insignificant angular anisotropy of 
sub--50--eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. However, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region. This effect is evident in 
Fig.~\ref{fig:emittedang};
for example, at 0.3 MeV/amu --where the Bragg peak for C$^{+6}$ impact 
occurs-- the CDW $\overline{\theta}_{\alpha}$ is half of the first Born 
value. This correction closes the damage to the forward direction.
We can attribute this forward direction correction to the capture to 
the continuum effect; the higher the charge $Z$, the smaller 
$\overline{\theta}$ will be. Of course, this effect only holds at 
intermediate energies and not at high impact energies.

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.6\textwidth}|}
\hline
\multirow{2}{*}{CH} 
      & CH$_4$ (methane), C$_2$H$_2$ (acetylene), C$_2$H$_4$ (ethene), \\
      & C$_2$H$_6$ (ethane), C$_6$H$_6$ (benzene) \\
\hline
\multirow{2}{*}{CHN} 
      & C$_5$H$_5$N (pyridine), C$_4$H$_4$N$_2$ (pyrimidine), \\ 
      & C$_2$H$_7$N (dimenthylamine), CH$_5$N (monomethylamine) \\
\hline
\multirow{4}{*}{DNA} 
      & C$_5$H$_5$N$_5$ (adenine), C$_4$H$_5$N$_3$O (cytosine), \\
      & C$_5$H$_5$N$_5$O (guanine), C$_5$H$_6$N$_2$O$_2$ (thymine), \\
      & C$_4$H$_4$N$_2$O$_2$ (uracil), C$_4$H$_8$O (THF), \\
      & C$_5$H$_{10}$O$_5$P (DNA backbone), 
        C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$ 
(dry DNA) \\
\hline
\end{tabular}
\caption{Molecular targets studied in this work, classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig2.eps}
\includegraphics[width=0.8\textwidth]{figuras/adn1.eps}
\caption{Reduced CDW ionization cross section by the impact of 
multicharged ions. Experiments: \mbox{\Large$\circ$}~\cite{iriki2011} 
for proton impact and $\square$~\cite{rahman2016} for electron impact 
with equivelocity conversion.}
\label{fig:crossDNA_1}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig3.eps}
\includegraphics[width=0.8\textwidth]{figuras/adn2.eps}
\caption{Reduced CDW ionization cross section by the impact of 
multicharged ions. Experiments: proton impact on 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$
THF~\cite{wang2016}. Impact of $\ominus$ C$^{+4}$, 
$\boxminus$ C$^{+6}$, O$^{+6}$, F$^{+6}$, and
$\triangle\mkern-14mu-$ O$^{+8}$, F$^{+8}$ on 
uracil~\cite{agnihotri2012,agnihotri2013}. 
Symbols~$\rhd$~\cite{bug2017}, $\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} for electron impact with equivelocity 
conversion.}
\label{fig:crossDNA_2}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig4.eps}
\includegraphics[width=0.8\textwidth]{figuras/molscaling85.eps}
\caption{Scaled ionization cross section per weakly bound electron using
(a)~the Toburen numbers $\nu_{\alpha}^T$, and (b) our proposed numbers
$\nu_{\alpha}^{\text{CDW}}$. Experiments: proton impact on 
\mbox{\Large$\circ$} adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$ 
THF~\cite{wang2016}; electron impact on $\rhd$ pyrimidine~\cite{bug2017},
and $\lhd$, $\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:newscaling}
\end{figure*}

@@

Based on the CDW results, we propose a new scaling,
\begin{equation}
\sigma_{e}=\frac{\sigma_M}{n_e'},
\label{32} 
\end{equation}
where $n_e'=\sum_{\alpha}n_{\alpha}\nu_{\alpha}^{\text{CDW}}$, and 
$\nu_{\alpha}^{\text{CDW}}$ are the numbers of active electrons
per atom obtained from the CDW ionization cross sections for 
different ions in H, C, N, O, P, and S targets, given as follows,
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
1, & \text{for H,} \\
4, & \text{for C, N, and O,} \\ 
4.5, & \text{for P and S}\,.
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}

The new scaled cross sections $\sigma_{e}$ are plotted in 
Fig.~\ref{fig:newscaling}b. 
A much better sharp band is observed, especially for impact energies 
$E=(0.5-8)$~MeV/amu for $Z=1$ and $Z=2$, and $E=(2.5-8)$~MeV/amu for 
$Z>2$. The experimental data for ionization of 
adenine~\cite{iriki2011}, uracil~\cite{itoh2013}, 
pyrimidine~\cite{wolff2014}, and THF~\cite{wang2016} by proton impact in
Fig.~\ref{fig:newscaling}b seems to corroborate the new scaling. 
We also included the electron impact ionization measurements with 
equivelocity conversion on pyrimidine~\cite{bug2017} and 
THF~\cite{bug2017,wolf2019,fuss2009}. 
It will be interesting to cross--check with future experiments, mainly 
for higher projectile charge states.

\begin{table}[H]
\begin{center}
\begin{tabular}{|ll|ll|ll|}
\hline
 Molecule & $n_e$ &Molecule          & $n_e$ & Molecule             & $n_e$\\
\hline
 H$_2$    & 2  & C$_2$H$_7$N         & 19    & C$_4$H$_5$N$_3$O     & 37   \\
 H$_2$O   & 6  & C$_4$H$_8$O         & 28    & C$_5$H$_6$N$_2$O$_2$ & 42   \\
 NH$_3$   & 7  & C$_4$H$_4$N$_2$     & 28    & C$_5$H$_5$N$_5$      & 45   \\
 CH$_4$   & 8  & C$_6$H$_6$          & 30    & C$_5$H$_5$N$_5$O     & 49   \\
 CH$_5$N  & 13 & C$_4$H$_4$N$_2$O$_2$& 36    & C$_5$H$_{10}$O$_5$P  & 54.5 \\
 \hline
\end{tabular}
\caption{New scaling numbers for some molecular targets of biological 
interest.}
\label{nn}
\end{center}
\end{table}

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig_recta.eps}
\includegraphics[width=0.8\textwidth]{figuras/scale_ne.eps}
\caption{Ionization cross sections by the impact of protons at 0.5, 1,
and 2 MeV in terms of the number of active electrons given by Table~\ref{nn}.
Experiments: 
\mbox{\Large$\circ$}~adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014}, 
$\blacktriangle$ C2H$_7$N, CH$_5$N, methane and ammonia~\cite{lynch1976},
\mbox{\scriptsize$\bigstar$} ammonia and H$_2$~\cite{rudd1985}, and 
\mbox{\Large$\bullet$} water~\cite{luna2007}.}
\label{fig:recta}
\end{figure*}

By using Eq.~(\ref{eq:scalingCDW}), we define new active electron 
numbers $n_e'$ for some molecules of interest in Table~\ref{nn}. These 
values are very different from the ones proposed by Toburen and used by 
other authors~\cite{itoh2013}. Moreover, an alternative way of showing 
the scaling can be attained by plotting the ionization cross sections 
of molecules as a function of the number of active electrons from 
Table~\ref{nn}. Our findings are displayed in Fig.~\ref{fig:recta} for 
impact energies 0.5, 1, and 2 MeV. As can be noted, the computed CDW 
ionization cross sections for all the molecules show a linear 
dependence with the number of electrons from Table~\ref{nn}. We obtain 
similar results even for $E=10$~MeV. The comparison with the 
experimental data available shows overall good agreement, for the 
smallest molecules, H$_2$, H$_2$O, and CH$_4$, up to the most complex 
ones, like adenine. For electron impact data, the experimental data was 
interpolated between close neighboors.

While finishing the present work, we noted an accepted manuscript by 
Luedde~{\it et al.} \cite{luedde2019} on total ionization of biological
molecules by proton impact, using an independent--atom--model pixel counting
method. The authors also raised a scaling with $\nu_{\alpha}=4$ 
for C, N, and O, but $\nu_{\alpha}=6$ for P. The agreement with this 
independent method for proton impact reinforces our multicharged--ion findings.

@@

To test the range of validity of the SSM, we performed a full molecular 
structure calculation of five nucleobases. We employed the {\sc gamess} 
code and used the 6-31G(d,p) basis set, which includes polarization 
functions for all the atoms. The calculations were carried out 
implementing the B3LYP functional~\cite{Becke1993,Stephens1994} to 
account for the correlation and exchange effects.

>>>>

%\documentclass[pra,showpacs,twocolumn]{revtex4}
%\documentclass[pra,showpacs]{revtex4}
%\documentclass[preprint,showpacs]{revtex4}
\documentclass[10pt,showpacs,twocolumn]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\title[Ionization of biological molecules by multicharged 
ions]{Ionization of biological molecules by multicharged ions using 
the stoichiometric model}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\address{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET--UBA), \\ Buenos Aires, Argentina.}

%\date{\today}% It is always \today, today,

\begin{abstract}
In the present work, we investigate the ionization of molecules of
biological interest by the impact of multicharged ions in the
intermediate to high energy range. We performed full non--perturbative
distorted--wave calculations (CDW) for thirty--six collisional systems
composed by six atomic targets: H, C, N, O, F and S --which
are the constituents of most of the DNA and biological molecules--, and
six charged projectiles (antiprotons, H, He, B, C, and O). On account
of the radiation damage caused by secondary electrons, we inspect
the energy and angular distributions of the emitted electrons.
We examine seventeen molecules: DNA and RNA bases,
DNA backbone, pyrimidines, tetrahydrofuran (THF), and several
C$_n$H$_n$ compounds. We show that the simple stoichiometric model
(SSM), which approximates the molecular ionization cross sections as a
linear combination of the atomic ones, gives reasonably good results
for complex molecules. We inspect the extensively used Toburen
scaling of the total ionization cross sections of molecules with the
number of weakly bound electrons. Based on the present atomic CDW
results, we propose new active electron numbers, which
improves significantly scaling for all the targets
and ions studied here.
The new scaling describes well the available experimental
data for proton impact, including small molecules such as H$_2$, water,
methane, and ammonia.
We perform full molecular
calculations for five nucleobases and test a modified stoichiometric
formula based on the Mulliken charge of the composite atoms. The
difference introduced by the new stoichiometric formula is less than
$3\%$, which indicates the reliability of the SSM to
deal with this type of molecules. The results of the extensive
ion--target examination included in the present study allow us to
assert that the SSM and the CDW--based scaling will be useful tools in
this area.
\end{abstract}

%\keywords{magnetic moment, solar neutrinos, astrophysics}
\pacs{34.50Gb, 34.80Gs, 34.80Dp}

\maketitle
%\ioptwocol

@@

In Section~\ref{subsec:stoichiometric}, we deal with the molecular 
structure complexity of the targets by 
implementing the simplest stoichiometric model (SSM): the molecules are 
assumed to be composed of isolated independent atoms, and the total 
cross section by a linear combination of stoichiometric weighted atomic 
calculations.
By implementing the CDW and the SSM, we calculate ionization 
cross section of several molecules of biological interest, including 
DNA and RNA molecules, such as adenine, cytosine, guanine, thymine, uracil, 
tetrahydrofuran~(THF), pyrimidine, and DNA backbone, by the impact 
of antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, and O$^{8+}$. 
In Section~\ref{subsec:scaling}, we
test the Toburen scaling rule~\cite{toburen1975,toburen1976}, which 
states that the ratio between the ionization cross section and the 
number of weakly bound electrons can be arranged in a narrow universal 
band in terms of the projectile velocity. We applied this rule to 
several hydrocarbons and nucleobases and noted that the width of the 
resulting universal band could be significantly reduced if we consider 
the number of active electrons in the collision based on the 
CDW results for the different atoms. The new scaling was then tested 
theoretically and by comparison with experimental data available.

@@

In the present study, we consider six atoms, $\alpha=$ H, C, N, O, P, 
and S, and six projectiles, antiprotons $\bar{p}$, H$^{+}$, He$^{2+}$, 
Be$^{4+}$, C$^{6+}$, and O$^{8+}$. 
Most of the organic molecules are composed of these atoms. Some 
particular molecules also include halogen atoms such as fluor and 
bromine; ionization cross sections of these elements have been 
previously published~\cite{miraglia2008}.

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{atomicscaling.eps}
\caption{Reduced CDW total ionization cross section of six atomic 
targets. The curves are labeled with the charge state corresponding to 
the six multicharged projectiles.}
\label{fig:atomscaling}
\end{figure*}

We display our total CDW ionization cross sections for the six essential 
elements by the impact of the six projectiles in Fig.~\ref{fig:atomscaling}.
To reduce the resulting 36 magnitudes into a single consistent 
figure, we considered the fact that in the first Born approximation
the ionization cross section scales with the square of the projectile 
charge, $Z^{2}$. The impact energies considered 
range between 0.1 to 10 MeV/amu, where the CDW is supposed 
to hold. In fact, for the highest projectile charges the minimum 
impact energy where the CDW is expected to be valid could be 
higher than 100 keV. We also performed similar calculations with the 
first Born approximation, and we corroborated that it provides quite 
reliable results only for energies higher than a couple of MeV/amu. 
We use the same line color to indicate the projectile charge throughout 
all the figures of this work: dashed--red, solid--red, blue, magenta, 
olive and orange for antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, 
C$^{6+}$, and O$^{8+}$, respectively. Notably, there is no complete 
tabulation of ionization of atoms by the impact of multicharged ions. 
We hope that the ones presented in this article will be of help for 
future works.

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{ener_mean.eps}
\caption{Mean emitted energy distribution for ionization by the impact 
of multicharged ions, given by Eq.~(\ref{40}). Dashed lines for $\bar{p}$,
solid lines for ion charges $+1$, $2+$, $4+$, $6+$ and $8+$,
as indicated.}
\label{fig:emittedener}
\end{figure*}

@@

In Fig.~\ref{fig:emittedener}, we estimate $\overline{E}_{\alpha}$ of
the emitted electron in the 10--70 eV energy range,
for all the targets. Our results agree with the 
experimental findings~\cite{surdutovic2018}. As can be noted in the 
figure, the mean energy value is surprisingly sensitive to the 
projectile charge $Z$, which can duplicate the proton results in the 
intermediate region, i.e., 100--400 keV/amu. The effect observed can be 
attributed to the depletion caused by the multicharged ions to the 
yields of low energy electrons. This behavior cannot be found in the 
first Born approximation, where the $Z^2$ law cancels the $Z$ dependence
in Eq.~(\ref{40}). At high energies, $\overline{E}_{\alpha}$ tends to a 
universal value for all ions, as can be seen in Fig.~\ref{fig:emittedener}.

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{ang_mean.eps}
\caption{Mean emitted angle distribution for ionization by impact of
multicharged ions.}
\label{fig:emittedang}
\end{figure*}

@@

@@

It is a general belief~\cite{Rudd1992} that the 
angular dispersion of emitted electrons is nearly isotropic. This 
behavior is thought to be caused by the insignificant angular anisotropy 
of sub--50--eV yield. 
A typical value for the ejection angle considered in the literature is 
$\overline{\theta}_{\alpha}\sim 70\degree$~\cite{surdutovic2018}, and 
it is quite correct in the range of validity of the first Born 
approximation for any target. However, when a distorted wave approximation 
is used, $\overline{\theta}_{\alpha}$ decreases substantially with $Z$ 
in the intermediate energy region. This effect is evident in 
Fig.~\ref{fig:emittedang};
for example, at 0.3 MeV/amu --where the Bragg peak for C$^{6+}$ impact 
occurs-- the CDW $\overline{\theta}_{\alpha}$ is half of the first Born 
value. This correction closes the damage to the forward direction.
We can attribute this forward direction correction to the capture to 
the continuum effect; the higher the charge $Z$, the smaller 
$\overline{\theta}$ will be. Of course, this effect only holds at 
intermediate energies and not at high impact energies.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|p{0.042\textwidth}|p{0.39\textwidth}|}
\hline
\multirow{2}{*}{CH} 
      & CH$_4$ (methane), C$_2$H$_2$ (acetylene),  \\
      & C$_2$H$_4$ (ethene), C$_2$H$_6$ (ethane), \\
      & C$_6$H$_6$ (benzene) \\
\hline
\multirow{2}{*}{CHN} 
      & C$_5$H$_5$N (pyridine), C$_4$H$_4$N$_2$ (pyrimidine), \\ 
      & C$_2$H$_7$N (dimenthylamine), \\
      & CH$_5$N (monomethylamine) \\
\hline
\multirow{4}{*}{DNA} 
      & C$_5$H$_5$N$_5$ (adenine), C$_4$H$_5$N$_3$O (cytosine), \\
      & C$_5$H$_5$N$_5$O (guanine), C$_5$H$_6$N$_2$O$_2$ (thymine), \\
      & C$_4$H$_4$N$_2$O$_2$ (uracil), C$_4$H$_8$O (THF), \\
      & C$_5$H$_{10}$O$_5$P (DNA backbone), \\
      & C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$ 
(dry DNA) \\
\hline
\end{tabular}
\caption{Molecular targets studied in this work, classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

@@

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{adn1.eps}
\caption{Reduced CDW ionization cross section by the impact of 
multicharged ions. Experiments: \mbox{\Large$\circ$}~\cite{iriki2011} 
for proton impact and $\square$~\cite{rahman2016} for electron impact 
with equivelocity conversion.}
\label{fig:crossDNA_1}
\end{figure*}

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{adn2.eps}
\caption{Reduced CDW ionization cross section by the impact of 
multicharged ions. Experiments: proton impact on 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$
THF~\cite{wang2016}. Impact of $\ominus$ C$^{4+}$, 
$\oplus$ C$^{6+}$, O$^{6+}$, F$^{6+}$, and
$\otimes$ O$^{8+}$, F$^{8+}$ on 
uracil~\cite{agnihotri2012,agnihotri2013}. 
Symbols~$\rhd$~\cite{bug2017}, $\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} for electron impact with equivelocity 
conversion.}
\label{fig:crossDNA_2}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{molscaling85.eps}
\caption{Scaled ionization cross section per weakly bound electron using
(a)~the Toburen numbers $\nu_{\alpha}^T$, and (b) our proposed numbers
$\nu_{\alpha}^{\text{CDW}}$. Experiments: proton impact on 
\mbox{\Large$\circ$} adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$ 
THF~\cite{wang2016}; electron impact on $\rhd$ pyrimidine~\cite{bug2017},
and $\lhd$, $\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:newscaling}
\end{figure*}

@@

Based on the CDW results, we propose a new scaling,
\begin{equation}
\sigma_{e}'=\frac{\sigma_M}{n_e'},
\label{32} 
\end{equation}
where $n_e'=\sum_{\alpha}n_{\alpha}\nu_{\alpha}^{\text{CDW}}$, and 
$\nu_{\alpha}^{\text{CDW}}$ are the numbers of active electrons
per atom obtained from the CDW ionization cross sections for 
different ions in H, C, N, O, P, and S targets, given as follows,
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
1, & \text{for H,} \\
4, & \text{for C, N, and O,} \\ 
4.5, & \text{for P and S}\,.
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}

The new scaled cross sections $\sigma_{e}'$ are plotted in 
Fig.~\ref{fig:newscaling}b. The experimental data for ionization of 
adenine~\cite{iriki2011}, uracil~\cite{itoh2013}, 
pyrimidine~\cite{wolff2014}, and THF~\cite{wang2016} by proton impact in
Fig.~\ref{fig:newscaling}b seems to corroborate the new scaling. 
We also included the electron impact ionization measurements with 
equivelocity conversion on pyrimidine~\cite{bug2017} and 
THF~\cite{bug2017,wolf2019,fuss2009}. 
It will be interesting to cross--check with future experiments, mainly 
for higher projectile charge states.

\begin{table}[t]
\begin{center}
\begin{tabular}{|p{0.07\textwidth}p{0.02\textwidth}|
p{0.095\textwidth}p{0.02\textwidth}|p{0.095\textwidth}p{0.035\textwidth}|}
\hline
 Molecule & $n_e'$ &Molecule          & $n_e'$ & Molecule             & $n_e'$\\
\hline
 H$_2$    & 2  & C$_2$H$_7$N         & 19    & C$_4$H$_5$N$_3$O     & 37   \\
 H$_2$O   & 6  & C$_4$H$_8$O         & 28    & C$_5$H$_6$N$_2$O$_2$ & 42   \\
 NH$_3$   & 7  & C$_4$H$_4$N$_2$     & 28    & C$_5$H$_5$N$_5$      & 45   \\
 CH$_4$   & 8  & C$_6$H$_6$          & 30    & C$_5$H$_5$N$_5$O     & 49   \\
 CH$_5$N  & 13 & C$_4$H$_4$N$_2$O$_2$& 36    & C$_5$H$_{10}$O$_5$P  & 54.5 \\
 \hline
\end{tabular}
\caption{New scaling numbers for some molecular targets of biological 
interest.}
\label{nn}
\end{center}
\end{table}

\begin{figure*}[t!]
\centering
\includegraphics[width=0.785\textwidth]{scale_ne.eps}
\caption{Ionization cross sections by the impact of protons at 0.5, 1,
and 2 MeV in terms of the number of active electrons given by Table~\ref{nn}.
Experiments: 
\mbox{\Large$\circ$}~adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014}, 
$\blacktriangle$ C2H$_7$N, CH$_5$N, methane and ammonia~\cite{lynch1976},
\mbox{\scriptsize$\bigstar$} ammonia and H$_2$~\cite{rudd1985}, and 
\mbox{\Large$\bullet$} water~\cite{luna2007}.}
\label{fig:recta}
\end{figure*}

By using Eq.~(\ref{eq:scalingCDW}), we define new active electron 
numbers $n_e'$ for some molecules of interest in Table~\ref{nn}. These 
values are different from the ones proposed by Toburen and used by 
other authors~\cite{itoh2013}. Moreover, an alternative way of showing 
the scaling can be attained by plotting the ionization cross sections 
of molecules as a function of the number of active electrons from 
Table~\ref{nn}. Our findings are displayed in Fig.~\ref{fig:recta} for 
impact energies 0.5, 1, and 2 MeV. As can be noted, the computed CDW 
ionization cross sections for all the molecules show a linear 
dependence with the number of electrons from Table~\ref{nn}. We obtain 
similar results, even for $E=10$~MeV. The comparison with the 
experimental data available shows overall good agreement, for the 
smallest molecules, H$_2$, H$_2$O, and CH$_4$, up to the most complex 
ones, like adenine. For electron impact data, the experimental data was 
interpolated between close neighbors.

While finishing the present work, we became aware of an accepted manuscript by 
Luedde~{\it et al.}~\cite{luedde2019} on total ionization of biological
molecules by proton impact, using an independent--atom--model pixel counting
method. The authors also raised a scaling with $\nu_{\alpha}=4$ 
for C, N, and O, but $\nu_{\alpha}=6$ for P. The agreement with this 
independent method for proton impact reinforces our multicharged--ion findings.

@@

To test the range of validity of the SSM, we performed {\it ab initio}
molecular calculation of five nucleobases by employing the {\sc gamess} 
code. The geometry optimization and single point energy calculations 
were performed implementing the restricted Hartree--Fock method and the 
3-21G basis set.

\begin{figure*}[t]
\centering
\includegraphics[width=0.8\textwidth]{levelsDNA.eps}
\caption{Theoretical molecular binding energies for adenine, cytosine, 
guanine, thymine, and uracil compared to those of atomic constituents.}
\label{fig:bindener}
\end{figure*}
